---
title: "Bootstrap Homework Assignment"
subtitle: "Raven Ico"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library(tidyverse))

```

# Example (Boos \& Stefanski Ex. 11.3)

The following data are from an experiment on food consumption of female rats treated with zinc-calcium EDTA are taken from Browne and Brownie (1986). The ordered sample values are: 

```{r}
rats = c(5.35, 5.37, 5.53, 5.95, 6.20, 7.12, 7.22,
         7.62, 7.63, 7.63, 7.67, 7.97, 8.43, 8.68,
         9.20, 9.63, 11.32, 11.52, 15.27, 15.90)
```

Suppose we are interested in estimating a 95\% confidence interval for
the median of food consumption in this group of rats. We can calculate this using the bootstrap as follows: 

```{r}
#**************************************************#
# Setup for the procedure
#**************************************************#
# set a seed for reproducibility
set.seed(42)
# set the number of bootstrap iterations
B = 1000 
# create a vector to store the statistics of interest
boot_medians = rep(NA, B)

#**************************************************#
# Main loop to do the calculations
#**************************************************#
for(i in 1:B)
{
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 1: resample data with replacement
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  this_rats = sample(x = rats, size = length(rats), replace = TRUE)
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 2: calculate the statistic and store the results
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # In this step our estimator is the median, but we can do a more 
  # complicated procedure than just using the median function
  boot_medians[i] = median(this_rats)
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 3: Repeat B times
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
}

# plot the results
hist(boot_medians)

#**********************************************************************#
# Calculate a 95% confidence interval from the vector of statistics
#**********************************************************************#
quantile(boot_medians, probs = c(0.025, 0.975))
```

Our 95\% confidence interval for the median is (7.17, 8.94). 


# Problem 1

Modify the code above to calculate a 95\% confidence interval for 
the inter-quartile range of the data (75th percentile - 25th percentile).
```{r}
#**************************************************#
# Setup for the procedure
#**************************************************#
# set a seed for reproducibility
set.seed(42)
# set the number of bootstrap iterations
B = 1000 
# create a vector to store the statistics of interest
boot_IQR = rep(NA, B)

#**************************************************#
# Main loop to do the calculations
#**************************************************#
for(i in 1:B)
{
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 1: resample data with replacement
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  this_rats = sample(x = rats, size = length(rats), replace = TRUE)
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 2: calculate the statistic (inter-quartile range) and store the results
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  boot_IQR[i] = IQR(this_rats)
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 3: Repeat B times
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
}

# plot the results
hist(boot_IQR, col="#CC79A7",main="Histogram of the simulated inter-quartile range (IQR) of rats data", xlab="IQR")

#**********************************************************************#
# Calculate a 95% confidence interval from the vector of statistics
#**********************************************************************#
tibble("Lower CI(2.5th quantile)"= quantile(boot_medians, probs = 0.025),"Lower CI(97.5th quantile)"= quantile(boot_medians, probs = 0.975))
```


# Problem 2

Use the bootstrap to calculated a 90\% confidence interval for the standard deviation of the rats data
```{r}
#**************************************************#
# Setup for the procedure
#**************************************************#
# set a seed for reproducibility
set.seed(42)
# set the number of bootstrap iterations
B = 1000 
# create a vector to store the statistics of interest
boot_sd = rep(NA, B)

#**************************************************#
# Main loop to do the calculations
#**************************************************#
for(i in 1:B)
{
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 1: resample data with replacement
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  this_rats = sample(x = rats, size = length(rats), replace = TRUE)
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 2: calculate the statistic (standard deviation) and store the results
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  boot_sd[i] = sd(this_rats)
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 3: Repeat B times
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
}

# plot the results
hist(boot_sd,col = "#E69F00",main="Histogram of the simulated standard deviation of the rats data",xlab="Standard Deviation")

#**********************************************************************#
# Calculate a 90% confidence interval from the vector of statistics
#**********************************************************************#
tibble("Lower CI(5th quantile)"= quantile(boot_sd, probs = 0.05),"Lower CI(95th quantile)"= quantile(boot_sd, probs = 0.95))
```

# Problem 3

Suppose our interest lies in calculating the difference in the median between two populations, x1 and x2. Using the bootstrap calculate a 95% confidence interval for this difference using the data below. 
(Hint: In Step 1 you should resample x1 and x2 separately)
```{r}
set.seed(42)
n1 = 60; n2 = 40; 
x1 = rgamma(n1, 2, 0.2)
x2 = rgamma(n2, 3, 0.5)

#**************************************************#
# Setup for the procedure
#**************************************************#
# set a seed for reproducibility
set.seed(42)
# set the number of bootstrap iterations
B = 1000 
# create a vector to store the statistics of interest
boot_median_diff = rep(NA, B)

#**************************************************#
# Main loop to do the calculations
#**************************************************#
for(i in 1:B)
{
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 1: resample data with replacement
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  resample_x1 = sample(x = x1, size = n1, replace = TRUE)
  resample_x2 = sample(x = x2, size = n2, replace = TRUE)
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 2: calculate the statistic(difference in median) and store the results
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  boot_median_diff[i] = median(resample_x1) - median(resample_x2)
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 3: Repeat B times
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
}

# plot the results
hist(boot_median_diff,col="#56B4E9",main="Histogram of the simulated difference in median of x1 and x2", xlab="Difference in median of x1 and x2")

#**********************************************************************#
# Calculate a 95% confidence interval from the vector of statistics
#**********************************************************************#
tibble("Lower CI(2.5th quantile)"= quantile(boot_median_diff, probs = 0.025),"Lower CI(97.5th quantile)"= quantile(boot_median_diff, probs = 0.975))
```

# Problem 4

We can also create confidence intervals for maximum likelihood estimators. 
Like in the MLE section, lets calculate a 95\% CI for $\lambda$ when we assume the data is from a Poisson distribution, this time using the bootstrap. 

```{r}
# use this as the data
x = c(3, 10, 2, 4, 8)
#**************************************************#
# Setup for the procedure
#**************************************************#
# set a seed for reproducibility
set.seed(42)
# set the number of bootstrap iterations
B = 1000 
# create a vector to store the statistics of interest
boot_lambda = rep(NA, B)

#**************************************************#
# Main loop to do the calculations
#**************************************************#
for(i in 1:B)
{
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 1: resample data with replacement
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  resample = sample(x = x, size = length(x), replace = TRUE)
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 2: calculate the statistic and store the results
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # In this step our estimator is the median, but we can do a more 
  # complicated procedure than just using the median function
  boot_lambda[i] = mean(resample)
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
  # Step 3: Repeat B times
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
}

# plot the results
hist(boot_lambda, col="#009E73", main="Histogram of the simulated value of lambda for x", xlab="Lambda")

#**********************************************************************#
# Calculate a 95% confidence interval from the vector of statistics
#**********************************************************************#
tibble("Lower CI(2.5th quantile)"= quantile(boot_lambda, probs = 0.025),"Lower CI(97.5th quantile)"= quantile(boot_lambda, probs = 0.975))

```

Compare this result with the CI calculated using aymptotic MLE theory
```{r}
n=length(x)

lower <- mean(x) - 2*sqrt(var(x)/n) #Theoretical lower and upper bounds for a 95% CI
upper <- mean(x) + 2*sqrt(var(x)/n)

tibble("Lower CI(Asymptotic MLE Theory)"= mean(x) - 2*sqrt(var(x)/n),
       "Upper CI(Asymptotic MLE Theory)"= mean(x) + 2*sqrt(var(x)/n))


```
**The obtained confidence intervals using the asymptotic MLE theory and bootstrap method are fairly close with one another.**