---
title: "Multiple Testing Lab"
subtitle: "Raven Ico"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library(tidyverse))

```

Use the following code to simulate a matric of categorical predictors $X \in \{0,1\}^{200\times100}$. Each column of $X$ represents a predictor and each row of $X$ indicates an observation. The code also simulates a response variable $Y$. 

```{r}
set.seed(2019)
# Simulate predictor matrix
X <- matrix(rbinom(10000, 1, .5), 200, 100) #100 predictors, 200 observations

# Simluate response vector
Y <- rnorm(200, mean = 2*X[,4]+5*X[,32]-.6*X[,90], sd = 1)
```

## Problem 1

What is the "true" effect of each of the predictors on $Y$ (we know this because we simulated the data)?

**True effect of X[,4] is 2, X[,32] is 5, and X[,90] is 0.6 and the other predictor variables is 0.**

## Problem 2
Write a for loop that conducts a hypothesis test comparing whether the mean of the values of $Y$ when the $i$th column of $X$ is $0$ are different than the mean of the values of $Y$ when the $i$th column of $X$ is $1$ for each of the simulated predictors. Store the p-values. How many p-values are less than 0.05?

```{r}
set.seed(2019)
pValues <- rep(NA, 100)
for(i in 1:100){
  pValues[i]<-t.test(Y~X[,i])$p.value
}
tibble("P-values<0.05 (No adjustment)"=sum(pValues<0.05))

```

## Problem 3
Use an appropriate method to control the familywise error rate. How many corrected p-values are less than 0.05? Using 0.05 as the cutoff how many false positives are there? How many true positives?

```{r}
Positives_Bonf = which(p.adjust(pValues,method="bonferroni")<0.05) #Correction method for familywise error rate
True_Positive_Bonf = sum(Positives_Bonf %in% c(4,32,90)) #True Positive
False_Positive_Bonf = length(Positives_Bonf)-True_Positive_Bonf

tibble("True Positive (Bonferroni)"=True_Positive_Bonf,"False Positive (Bonferroni)"=False_Positive_Bonf, "Total Positives (corrected p-values<0.05)" = length(Positives_Bonf))

```

## Problem 4
Use an appropriate method to control the false discovery rate. How many corrected p-values are less than 0.05? Using 0.05 as the cutoff how many false positives are there? How many true positives?

```{r}
#Method of correction for False Discovery Rate = Benjamini & Hochberg
Positives_BH = which(p.adjust(pValues,method="BH")<0.05) #Correction method for familywise error rate
True_Positive_BH = sum(Positives_BH %in% c(4,32,90)) #True Positive
False_Positive_BH = length(Positives_BH)-True_Positive_BH

tibble("True Positive (BH)"=True_Positive_BH,"False Positive (BH)"=False_Positive_BH, "Total Positives (corrected p-values<0.05)" = length(Positives_BH))

```

