---
title: "R Notes"
author: "Raven Ico"
date: "September 21, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(UsingR)
data(galton)
library(reshape2)
long <- melt(galton)
ggplot(long,aes(x=value, fill=variable)) +
  geom_histogram(colour = "black", binwidth = 1) + facet_grid(.~variable)
```
---
## Experiment
### Use R studio's manipulate to see what value of $\mu$ minimizes the sum of the squared deviations.

```{r}
library(manipulate)
myHist <- function(mu){
    mse <- mean((galton$child - mu)^2)
    g <- ggplot(galton, aes(x = child)) + geom_histogram(fill = "salmon", colour = "black", binwidth=1)
    g <- g + geom_vline(xintercept = mu, size = 3)
    g <- g + ggtitle(paste("mu = ", mu, ", MSE = ", round(mse, 2), sep = ""))
    g
}
manipulate(myHist(mu), mu = slider(62, 74, step = 0.5))
```
Size of point represents number of points at that (X, Y) combination (See the Rmd file for the code).
```{r}
library(dplyr)
freqData <- as.data.frame(table(galton$child, galton$parent))
names(freqData) <- c("child", "parent", "freq")
freqData$child <- as.numeric(as.character(freqData$child))
freqData$parent <- as.numeric(as.character(freqData$parent))
g <- ggplot(filter(freqData, freq > 0), aes(x = parent, y = child))
g <- g  + scale_size(range = c(2, 20), guide = "none" )
g <- g + geom_point(colour="grey50", aes(size = freq+20, show_guide = FALSE))
g <- g + geom_point(aes(colour=freq, size = freq))
g <- g + scale_colour_gradient(low = "lightblue", high="white")                    
g
```
Regression through the origin
```{r}
y <- galton$child - mean(galton$child)
x <- galton$parent - mean(galton$parent)
freqData <- as.data.frame(table(x, y))
names(freqData) <- c("child", "parent", "freq")
freqData$child <- as.numeric(as.character(freqData$child))
freqData$parent <- as.numeric(as.character(freqData$parent))
myPlot <- function(beta){
    g <- ggplot(filter(freqData, freq > 0), aes(x = parent, y = child))
    g <- g  + scale_size(range = c(2, 20), guide = "none" )
    g <- g + geom_point(colour="grey50", aes(size = freq+20, show_guide = FALSE))
    g <- g + geom_point(aes(colour=freq, size = freq))
    g <- g + scale_colour_gradient(low = "lightblue", high="white")                     
    g <- g + geom_abline(intercept = 0, slope = beta, size = 3)
    mse <- mean( (y - beta * x) ^2 )
    g <- g + ggtitle(paste("beta = ", beta, "mse = ", round(mse, 3)))
    g
}
manipulate(myPlot(beta), beta = slider(0.6, 1.2, step = 0.02))
```
Solution
```{r}
lm(I(child - mean(child))~ I(parent - mean(parent)) - 1, data = galton)
```

---
```{r, fig.height=6,fig.width=7,echo=FALSE}
freqData <- as.data.frame(table(galton$child, galton$parent))
names(freqData) <- c("child", "parent", "freq")
freqData$child <- as.numeric(as.character(freqData$child))
freqData$parent <- as.numeric(as.character(freqData$parent))
g <- ggplot(filter(freqData, freq > 0), aes(x = parent, y = child))
g <- g  + scale_size(range = c(2, 20), guide = "none" )
g <- g + geom_point(colour="grey50", aes(size = freq+20, show_guide = FALSE))
g <- g + geom_point(aes(colour=freq, size = freq))
g <- g + scale_colour_gradient(low = "lightblue", high="white")                    
lm1 <- lm(galton$child ~ galton$parent)
g <- g + geom_abline(intercept = coef(lm1)[1], slope = coef(lm1)[2], size = 3, colour = grey(.5))
g
```
Statistical Linear Regression
```{r}
library(UsingR)
data(diamond)
library(ggplot2)
g = ggplot(diamond, aes(x = carat, y = price))
g = g + xlab("Mass (carats)")
g = g + ylab("Price (SIN $)")
g = g + geom_point(size = 7, colour = "black", alpha=0.5)
g = g + geom_point(size = 5, colour = "blue", alpha=0.2)
g = g + geom_smooth(method = "lm", colour = "black")
g
```
Fitting the linear regression model
```{r}
#price is the response, carat is the predictor
fit <- lm(price ~ carat, data = diamond)
coef(fit)
#summary(fit)
```
Getting a more interpretable intercept
```{r}
fit2 <- lm(price ~ I(carat - mean(carat)), data = diamond)
coef(fit2)
#the slope is the same, 500.1 is the expected size of the average sized diamond
```
Changing scale
* A one carat increase in a diamond is pretty big, what about
  changing units to 1/10th of a carat? 
* We can just do this by just dividing the coeficient by 10.
  * We expect  a `r round(coef(fit)[2], 2) / 10` (SIN) dollar   change in price for every 1/10th of a carat increase in mass of diamond.
* Showing that it's the same if we rescale the Xs and refit
```{r}
fit3 <- lm(price ~ I(carat * 10), data = diamond)
coef(fit3)
```
Predicting the price of the diamond
```{r, echo = TRUE}
newx <- c(0.16, 0.27, 0.34)
coef(fit)[1] + coef(fit)[2] * newx
predict(fit, newdata = data.frame(carat = newx))
```
Predicted values at the observed Xs (red)
and at the new Xs (lines)
```{r}
data(diamond)
plot(diamond$carat, diamond$price,  
     xlab = "Mass (carats)", 
     ylab = "Price (SIN $)", 
     bg = "lightblue", 
     col = "black", cex = 1.1, pch = 21,frame = FALSE)
abline(fit, lwd = 2)
points(diamond$carat, predict(fit), pch = 19, col = "red")
lines(c(0.16, 0.16, 0.12), 
      c(200, coef(fit)[1] + coef(fit)[2] * 0.16,
      coef(fit)[1] + coef(fit)[2] * 0.16))
lines(c(0.27, 0.27, 0.12), 
      c(200, coef(fit)[1] + coef(fit)[2] * 0.27,
        coef(fit)[1] + coef(fit)[2] * 0.27))
lines(c(0.34, 0.34, 0.12), 
      c(200, coef(fit)[1] + coef(fit)[2] * 0.34,
        coef(fit)[1] + coef(fit)[2] * 0.34))
text(newx, rep(250, 3), labels = newx, pos = 2)
```
